---
title: "RF_Prep"
author: "Daniel Dominguez"
date: '2023-05-25'
output: html_document
editor_options: 
  chunk_output_type: console
---
### The best way to describe how the training framework works/ was designed is by thikning about car engine tuning for an automatic transmission. As the DOC (or water quality parameter) increases the AI model switches similar to what occurs when you step on the gas and the car switches gears. In this case each of the AI models can be thought about as a gear in the transmission. The statistics of another AI predicting a measurement in a different quanitle also back up how this works but will be explained below. 

```{r}
local_Path <- "/Users/danieldominguez/Documents/Code/ATS_RandomForests"

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(feather)
library(viridis)
library(sf)
library(mlbench)
library(caret)
library(xgboost)
library(Metrics)
library(purrr)
library(data.table)
library(mltools)
library(ggthemes)
library(dplyr)
library(ggplot2)
library(scales)
library(dplyr)
library(kableExtra)

```

```{r functions, echo=FALSE}

round_any = function(x, accuracy, f = round) {
  f(x / accuracy) * accuracy
}

traintest <- function(x) {
  # quantile is represented in each fold to make them representative. Basically takes a more stratified sample
  x <- x %>%
    dplyr::mutate(mag = cut(value, quantile(
      x = value,
      c(0,0.05,0.1, 0.2,0.3, 0.4,0.5, 0.6, 0.7,  0.8,  0.9,0.95,  1),
      include.lowest = T
    )),
    mag = factor(
      mag,
      labels = c( 'A0.05','A0.1', 'A0.2','A0.3', 'A0.4','A0.5', 'A0.6', 'A0.7',  'A0.8',  'A0.9','A0.95','A1')
    )) %>%
    ungroup()
  
  set.seed(22)
  
  train <- x %>%
    group_by( mag,ecoregion) %>%
    sample_frac(.8) %>%
    ungroup() %>%
    dplyr::mutate(.partitions = 1)
  
  test <- x %>%
    filter(!(uniqueID %in% train$uniqueID)) %>%
    dplyr::mutate(.partitions = 2)
  
  out <- train %>%
    bind_rows(test)
  
  return(out)
}

random_filter_data <- function(df, column_name, condition, filter_percentage, seed ) {
  # Set a random seed for reproducibility
  if (!is.null(seed)) {
    set.seed(seed)
  }
  
  # Create a logical vector indicating whether rows satisfy the condition
  mask <- sapply(df[[column_name]], condition)
  
  # Randomly select the indices to keep based on the filter_percentage
  num_to_keep <- round(sum(mask) * (1 - filter_percentage))
  indices_to_keep <- sample(which(mask), num_to_keep)
  
  # Filter the DataFrame based on the selected indices
  filtered_df <- df[indices_to_keep, ]
  
  return(filtered_df)
}

```


```{r prep}
# filter data
data <- read.csv(paste0(local_Path,"/data/aquasat_processed.csv")) %>%
    #filter(!is.na(doc)) %>% # remove rows without DOC data
  filter(!is.na(doc) & !is.na(secchi)) %>% 
    filter(pwater>=80) %>% #filter for mostly water pixels
    filter(!type=="Facility") %>% #not enough data to be representative when splitting betwen quanitles
    #filter(n>=10) %>% # filter if there is not at least ten readings for a site for better training
    filter(pixelCount >0 ) %>% #filter erroneous pixel counts
    mutate(value = doc) %>%  #mutate doc to log
  filter_all(all_vars(!is.infinite(.))) %>% # remove rows with infinites that might have resulted from band calculations
  mutate(WC = case_when(
    red < green & red > blue ~ "Green",
    red > green & nir > 0.01 ~ "Brown",
    green < blue ~ "BG",
    TRUE ~ "Green"
  ))
  
```

```{r}
bands <- c("red", "green","blue", "swir1", "swir2","nir",
  "NR", "BR", "GR", "SR", "BG", "RG", "NG", "SG", "BN", "GN",
  "RN", "SN", "BS", "GS", "RS", "NS", "RGN", "RGB", "RGS", "RBN",
  "RBS", "RNS", "GBR", "GBN", "GBS", "GRN", "GRB", "GNS",
  "BRG", "BRS", "BGN", "BGS", "BNS", "NRG", "NRB", "NRS",
  "NGB", "NGS", "NBS", "GR2", "GN2", "B_RG", "NS_NR", "fai",
  "N_S", "N_R", "NDVI", "NDWI", "NDSSI", "GNGN"
)
```


```{r}

# select featuers
features <- c("red", "blue","GS","green",
              "swir1","swir2","SG","RBS","NGS","GBS","GR2","GN2",
              "NR","N_S", "RGB","ecoregion", "type","season","WC")  

filtered_data <- data %>%
  droplevels() %>%
  traintest() %>% 
  ungroup() %>%
  dplyr::select(bands, value, uniqueID,ecoregion, type,season,mag,WC,secchi) %>% 
  filter(!is.na(mag)) %>%
  mutate_if(is.character,as.factor) 


filtered_data$mag <- ifelse(filtered_data$value <= 10, "low", "high")

rf_data_low<-random_filter_data(filtered_data, 'value', function(x) x > 0 & x < 10, filter_percentage = 0.95, seed = 01) %>% 
  dplyr::select(bands, value, uniqueID,ecoregion, type,season,WC,secchi) %>% 
  mutate(mag="low")

rf_data_med<-random_filter_data(filtered_data, 'value', function(x) x > 10 & x < 40, filter_percentage = 0.9, seed = 01) %>% 
  dplyr::select(bands, value, uniqueID,ecoregion, type,season,WC,secchi)%>% 
  mutate(mag="low")

rf_data_high<- data %>% 
  filter(value>=40) %>% 
  dplyr::select(bands, value, uniqueID,ecoregion, type,season,WC,secchi)%>% 
  mutate(mag="high")

rf_data<-rbind(rf_data_low,rf_data_med,rf_data_high)

  write.csv(filtered_data,paste0(local_Path,"/data/aquasat_full.csv"),row.names=FALSE)

write.csv(rf_data,paste0(local_Path,"/data/rf_aquasat_full.csv"),row.names=FALSE)  

```

```{r}
  train_v1 <- filtered_data %>%
    group_by( mag) %>%
    sample_frac(.8) %>%
    ungroup() %>%
  arrange(value)
  
  test_v1 <- filtered_data %>%
    filter(!(uniqueID %in% train_v1$uniqueID)) %>%
  arrange(value)
  
write.csv(train_v1,paste0(local_Path,"/data/train_v1.csv"),row.names=FALSE)

write.csv(test_v1,paste0(local_Path,"/data/test_v1.csv"),row.names=FALSE)
    
    
```

